### 2020-05-27 writeup/proof_examples.tex
restructure appendix;
prop 4.4

@@

Based on Lemma \ref{lem_hat_v}, we can prove Proposition \ref{prop_dist_transition}.
\begin{proof}[Proof of Proposition \ref{prop_dist_transition}]
	Since $\Sigma_1 = \Sigma_2$, we know that $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} = \hat{v} \id$.
	Using Lemma \ref{lem_minv} and \ref{lem_cov_shift}, we can track the reduction of variance from $\hat{\beta}_t^{\MTL}$ to $\hat{\beta}_t^{\STL}$ as whp
	\begin{align}
		\Delta_{\vari} &\define \sigma^2 \bigbrace{\frac{1}{c_2 - 1} - \frac{1}{c_1 + c_2}\cdot\frac{1}{a_1\hat{v}^2 + a_2} + \bigo{p^{-1/2+\varepsilon}}} \nonumber\\
		&= \sigma^2 \bigbrace{\frac{c_1}{(c_2-1)(c_1 + c_2 -1)} + \bigo{p^{-1/2 + \varepsilon}}},\label{Deltavar}
	\end{align}
	where we use equation \eqref{eq_a12} and Lemma \ref{lem_hat_v}.
	Next we consider the model shift bias
	\begin{align*}
		\Delta_{\beta} &\define \hat{v}^2 \bignorm{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v}\beta_t)}^2 \\
		&= d^2 \cdot \bignormFro{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1}^2 + \bigo{p^{-1/2 + \varepsilon}d^2}
	\end{align*}
	Using Lemma \ref{prop_model_shift_tight}, we get an upper and lower bound on $\Delta_{\beta}$ as
	\be\label{Deltabeta} \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 \le \Delta_{\beta} / \bigbrace{p\cdot d^2 \cdot \frac{c_1^2}{(c_1+c_2)^2} \cdot \frac{1 + a_3 + a_4}{(a_1 + a_2)^2} + \bigo{p^{1/2+\e}d^2} } \le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4. \ee
	By solving equations \eqref{eq_a12}, \eqref{eq_a3} and \eqref{eq_a4}, we get
	\begin{align*}
		& a_1 = \frac{c_1(c_1 + c_2 - 1)}{(c_1 + c_2)^2} + O(p^{-1/2 + \varepsilon}),
		a_2 = \frac{c_2(c_1 + c_2 - 1)}{(c_1 + c_2)^2} + O(p^{-1/2 + \varepsilon}), \\
		& a_3 = \frac{c_2}{(c_1 + c_2)(c_1 + c_2 - 1)} + O(p^{-1/2 + \varepsilon}),
		a_4 = \frac{c_1}{(c_1 + c_2)(c_1 + c_2 - 1)} + O(p^{-1/2 + \varepsilon}).
	\end{align*}
	Hence we obtain that
	\begin{align*}
		\frac{1 + a_3 + a_4}{(a_1 + a_2)^2}
		= \frac{(c_1 + c_2)^3}{(c_1 + c_2 - 1)^3} + \bigo{p^{-1/2+\varepsilon}}.
	\end{align*}
	To sum up, we have shown that when
	\[ \frac{p\cdot d^2}{\sigma^2} < \frac{(c_1 + c_2 - 1)^2}{c_1(c_1 + c_2) (c_2 - 1)} \cdot \bigbrace{1 + \sqrt{\frac 1 {c_1}}}^{-4}, \]
	we have that $\Delta_{\vari} > \Delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$.
	When
	\[ \frac{p\cdot d^2}{\sigma^2} > \frac{(c_1 + c_2 - 1)^2}{c_1 (c_1 + c_2) (c_2 - 1)} \cdot \bigbrace{1 - \sqrt{\frac 1 {c_1}}}^{-4}, \]
	we have that $\Delta_{\vari} < \Delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is higher than $\te(\hat{\beta}_t^{\STL})$.
	The result follows by taking $c_1 = c_2 = c$ for the above equation.
\end{proof}

@@

\textbf{Remark.} Furthermore, as a function of $c_1$ over the range $[??, \infty]$, the maximum of $\te(\hat{\beta}_t^{\STL}) - \te(\hat{\beta}_t^{\MTL})$ is attained when $c_1 = {c_2\sigma^2}/{\max(2(c_2 - 1)pd^2 -\sigma^2, 0)}$. {\cor (cannot get this because we only have some bounds. If we let $c_1\to \infty$, then the curve of $\te(\hat{\beta}_t^{\STL}) - \te(\hat{\beta}_t^{\MTL})$ already becomes flat, and it is meaningless to discuss the minimum of this function at this point?)}

\begin{proof}[Proof of Proposition \ref{prop_data_efficiency}]
Suppose we have reduced number of datapoints---$\al n_1$ for task 1 and $\al n_2$ for task 2 with $n_1=n_2$. In the setting of Proposition \ref{prop_data_efficiency}, we still have \eqref{hatw_add1}, and by symmetry it suffices to focus on one of tasks, say task 2. Using Lemmas \ref{lem_cov_shift} and \ref{prop_model_shift_tight}, we have that whp,
\be\label{reduceproof1}
\begin{split}
\te_2(\hat \beta(\al)) =\sigma^2\left( \frac{1}{2\al c - 1}+ \bigo{p^{-1/2 + \varepsilon}}\right)+ \Delta^{(2)}_{\beta},
\end{split}
\ee
where $ \Delta^{(2)}_{\beta}$ satisfies 
\be \nonumber
\bigbrace{1 - \sqrt{\frac{1}{\al c}}}^4 \le \Delta^{(2)}_{\beta} /  \left( pd^2 \cdot \frac{2\al^3 c^3}{(2\al c-1)^3}+ \bigo{p^{1/2+\e}d^2}\right) \le \bigbrace{1 + \sqrt{\frac{1}{\al c}}}^4. 
\ee
On the other hand, using Lemma \ref{lem_minv}, we have whp
\be\label{reduceproof2} 
\te_2(\hat{\beta}_t^{\STL}) = \frac{\sigma^2}{c-1} \left( 1+ \bigo{p^{-1/2 + \varepsilon}}\right).
\ee
Comparing \eqref{reduceproof1} and \eqref{reduceproof2}, we observe that $\te_2(\hat \beta(\al))  \ge \te_2(\hat{\beta}_t^{\STL}) $ for $\al \le 1/2-\oo(1)$, which gives $\al^* \ge 1/2 - \oo(1)$.

Now for $c\ge 200$ and $\al \ge 1/2$, we have whp
\be\label{reduceproof3} \te_2(\hat \beta(\al)) < \sigma^2 \frac{1}{2\al c - 1}+  \frac38 pd^2  \ee
for large enough $p$. Moreover, if 
$$\al \ge \al_0:=\frac{1}{2c} + \left(2-\frac{3}{4}c \cdot \frac{pd^2}{\sigma^2} \right)^{-1},$$
we can check from \eqref{reduceproof3} that $\te_2(\hat \beta(\al))  < \te_2(\hat{\beta}_t^{\STL}) $. By symmetry, we also have $\te_1(\hat \beta(\al))  < \te_1(\hat{\beta}_t^{\STL}) $. Thus we conclude that 
$$\te_1(\hat{\beta}(\alpha)) + \te_2(\hat{\beta}(\alpha))\le \te_1(\hat{\beta}_t^{\STL}) + \te_2(\hat{\beta}_t^{\STL})$$
for $\al\ge \al_0$, which shows that $\al^*\le \al_0$.
\end{proof}

>>>>

\section{Supplementary Materials for the Theoretical Implications}

@@

Based on Lemma \ref{lem_hat_v}, we can prove Proposition \ref{prop_dist_transition}.
\begin{proof}[Proof of Proposition \ref{prop_dist_transition}]
	Since $\Sigma_1 = \Sigma_2$, we know that $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} = \hat{v} \id$.
	Using Lemma \ref{lem_minv} and \ref{lem_cov_shift}, we can track the reduction of variance from $\hat{\beta}_t^{\MTL}$ to $\hat{\beta}_t^{\STL}$ as whp
	\begin{align}
		\Delta_{\vari} &\define \sigma^2 \bigbrace{\frac{1}{c_2 - 1} - \frac{1}{c_1 + c_2}\cdot\frac{1}{a_1\hat{v}^2 + a_2} + \bigo{p^{-1/2+\varepsilon}}} \nonumber\\
		&= \sigma^2 \bigbrace{\frac{c_1}{(c_2-1)(c_1 + c_2 -1)} + \bigo{p^{-1/2 + \varepsilon}}},\label{Deltavar}
	\end{align}
	where we use equation \eqref{eq_a12} and Lemma \ref{lem_hat_v}.
	Next we consider the model shift bias
	\begin{align*}
		\Delta_{\beta} &\define \hat{v}^2 \bignorm{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_s - \hat{v}\beta_t)}^2 \\
		&= d^2 \cdot \bignormFro{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1}^2 + \bigo{p^{-1/2 + \varepsilon}d^2}
	\end{align*}
	Using Lemma \ref{prop_model_shift_tight}, we get an upper and lower bound on $\Delta_{\beta}$ as
	\be\label{Deltabeta} \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 \le \Delta_{\beta} / \bigbrace{p\cdot d^2 \cdot \frac{c_1^2}{(c_1+c_2)^2} \cdot \frac{1 + a_3 + a_4}{(a_1 + a_2)^2} + \bigo{p^{1/2+\e}d^2} } \le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4. \ee
	By solving equations \eqref{eq_a12}, \eqref{eq_a3} and \eqref{eq_a4}, we get
	\begin{align}
		& a_1 = \frac{c_1(c_1 + c_2 - 1)}{(c_1 + c_2)^2} + O(p^{-1/2 + \varepsilon}),
		a_2 = \frac{c_2(c_1 + c_2 - 1)}{(c_1 + c_2)^2} + O(p^{-1/2 + \varepsilon}), \label{simplesovlea12}\\
		& a_3 = \frac{c_2}{(c_1 + c_2)(c_1 + c_2 - 1)} + O(p^{-1/2 + \varepsilon}),
		a_4 = \frac{c_1}{(c_1 + c_2)(c_1 + c_2 - 1)} + O(p^{-1/2 + \varepsilon}).\label{simplesovlea34}
	\end{align}
	Hence we obtain that
	\begin{align*}
		\frac{1 + a_3 + a_4}{(a_1 + a_2)^2}
		= \frac{(c_1 + c_2)^3}{(c_1 + c_2 - 1)^3} + \bigo{p^{-1/2+\varepsilon}}.
	\end{align*}
	To sum up, we have shown that when
	\[ \frac{p\cdot d^2}{\sigma^2} < \frac{(c_1 + c_2 - 1)^2}{c_1(c_1 + c_2) (c_2 - 1)} \cdot \bigbrace{1 + \sqrt{\frac 1 {c_1}}}^{-4}, \]
	we have that $\Delta_{\vari} > \Delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$.
	When
	\[ \frac{p\cdot d^2}{\sigma^2} > \frac{(c_1 + c_2 - 1)^2}{c_1 (c_1 + c_2) (c_2 - 1)} \cdot \bigbrace{1 - \sqrt{\frac 1 {c_1}}}^{-4}, \]
	we have that $\Delta_{\vari} < \Delta_{\beta}$, which implies that $\te(\hat{\beta}_t^{\MTL})$ is higher than $\te(\hat{\beta}_t^{\STL})$.
	The result follows by taking $c_1 = c_2 = c$ for the above equation.
\end{proof}

@@

%\textbf{Remark.} Furthermore, as a function of $c_1$ over the range $[??, \infty]$, the maximum of $\te(\hat{\beta}_t^{\STL}) - \te(\hat{\beta}_t^{\MTL})$ is attained when $c_1 = {c_2\sigma^2}/{\max(2(c_2 - 1)pd^2 -\sigma^2, 0)}$. {\cor (cannot get this because we only have some bounds. If we let $c_1\to \infty$, then the curve of $\te(\hat{\beta}_t^{\STL}) - \te(\hat{\beta}_t^{\MTL})$ already becomes flat, and it is meaningless to discuss the minimum of this function at this point?)}

\begin{proof}[Proof of Proposition \ref{prop_data_efficiency}]
Suppose we have reduced number of datapoints---$\al n_1$ for task 1 and $\al n_2$ for task 2 with $n_1=\rho_1 p$ and $n_2=\rho_2 p$. 
%For this proof, we introduce the notations 
%$$\al_{\pm}^{(1)}\equiv \al_{\pm}^{(1)}(\al) = \left ( 1- (\al\rho_1)^{-1/2}\right)^2,\quad \al_{\pm}^{(2)}\equiv \al_{\pm}^{(2)}(\al) = \left ( 1- (\al\rho_2)^{-1/2}\right)^2.$$
In the setting of Proposition \ref{prop_data_efficiency}, we still have \eqref{hatw_add1}. 
%and by symmetry it suffices to focus on one of tasks, say task 2. 
Then using Lemmas \ref{lem_cov_shift} and \ref{prop_model_shift_tight}, and recalling \eqref{simplesovlea12} and \eqref{simplesovlea34}, we get that whp,
\be\label{reduceproof1}
\begin{split}
\te_i(\hat \beta(\al)) =\sigma^2\left( \frac{1}{\al (\rho_1+\rho_2) - 1}+ \bigo{p^{-1/2 + \varepsilon}}\right)+ \Delta^{(i)}_{\beta}, \quad i=1,2,
\end{split}
\ee
%and
%\be\label{reduceproof1}
%\begin{split}
%\te_2(\hat \beta(\al)) =\sigma^2\left( \frac{1}{\al ( \rho_1+\rho_2) - 1}+ \Delta^{(2)}_{\beta}+ \bigo{p^{-1/2 + \varepsilon}}\right),
%\end{split}
%\ee
where $ \Delta^{(i)}_{\beta}$ satisfies 
\be \nonumber
\bigbrace{1 - \sqrt{\frac{1}{\al \rho_i}}}^4 \le \Delta^{(i)}_{\beta} /  \left[ pd^2 \cdot \frac{(\al\rho_i)^2\cdot \al (\rho_1+\rho_2)}{[\al ( \rho_1+\rho_2) - 1]^3}\cdot\left(1+ \bigo{p^{-1/2+\e}}\right)\right] \le \bigbrace{1 + \sqrt{\frac{1}{\al \rho_i}}}^4. 
\ee
On the other hand, using Lemma \ref{lem_minv}, we have whp
\be\label{reduceproof2} 
\te_i(\hat{\beta}_t^{\STL}) = \frac{\sigma^2}{\rho_i -1} \left( 1+ \bigo{p^{-1/2 + \varepsilon}}\right),\quad i=1,2.
\ee
Comparing \eqref{reduceproof1} and \eqref{reduceproof2}, we immediately obtain a trivial lower bound $\al^*\ge \al_l -\oo(1)$, where
$$\al_l:=\frac1{\rho_1+\rho_2}\left[\frac{2(\rho_1-1)(\rho_2-1)}{\rho_1+\rho_2 -2}+1\right] \ge \frac{\min(\rho_1,\rho_2)}{\rho_1+\rho_2}.$$
In fact, one can see that if $\al\le \al_l$, then we have 
$$ \frac{2\sigma^2}{\al (\rho_1+\rho_2) - 1} \ge \frac{\sigma^2}{\rho_1-1}+\frac{\sigma^2}{\rho_2-1},$$
that is $\te_1(\hat{\beta}(\alpha)) + \te_2(\hat{\beta}(\alpha))$ is larger than $\te_1(\hat{\beta}_t^{\STL}) + \te_2(\hat{\beta}_t^{\STL})$ even if we do not take into account the model shift bias terms $ \Delta^{(i)}_{\beta}$. Next we try to get more precise bounds on $\al^*$. In the following discussions, we only consider $\al$ such that $\al\rho \ge \al_l \rho \ge \min(\rho_1,\rho_2)$.

%Comparing \eqref{reduceproof1} and \eqref{reduceproof2}, we observe that $\te_2(\hat \beta(\al))  \ge \te_2(\hat{\beta}_t^{\STL}) $ for $\al \le 1/2-\oo(1)$, which gives $\al^* \ge 1/2 - \oo(1)$.

\noindent{\bf The upper bound.} From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $\al^*\le \al$ if $\al$ satisfies
\be\label{solval1}
\begin{split}
&(1+\oo(1)) \cdot \sum_{i=1}^2 pd^2 \frac{(\al\rho_i)^2\cdot \al (\rho_1+\rho_2)}{[\al ( \rho_1+\rho_2) - 1]^3}  \bigbrace{1 + \sqrt{\frac{1}{\al \rho_i}}}^4 \\
&\le \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{\al (\rho_1+\rho_2) - 1} . 
\end{split}
\ee
For the sum on the left-hand side, we can rewrite it as
\begin{align*}
 pd^2 \frac{1}{[1 - (\al \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho_1+\rho_2}} + \sqrt{\frac{1}{\al \rho}}}^4,
\end{align*}
where $\rho:=\rho_1+\rho_2$.

In order to solve \eqref{solval1}, we now consider the case $\min(\rho_1,\rho_2)\ge 200$. With some basic calculations, one can show that in this case
$$ \frac{1}{[1 - (\al \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho_1+\rho_2}} + \sqrt{\frac{1}{\al \rho}}}^4 <  1.32.$$%\frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + 0.32.$$
Thus the following inequality implies \eqref{solval1}:
\be\label{solval1add}
\begin{split}
&1.32 pd^2 < \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{\al (\rho_1+\rho_2) - 1} .
\end{split}
\ee
In particular, if 
\be\nonumber
\begin{split}
&1.32 pd^2 < \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{(\rho_1+\rho_2) - 1} ,
\end{split}
\ee
that is, we have positive transfer when using all the data, then we can solve from \eqref{solval1add} the following upper bound on $\al^*$:
\begin{align*}
\al^* &<  \frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  - (\rho_1-1)(\rho_2-1) \cdot 1.32\frac{pd^2}{\sigma^2}}+ 1\right] \\
& < \frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1}+\frac1{\rho_2}  - \frac43\frac{pd^2}{\sigma^2}}+ 1\right] .
\end{align*}

\noindent{\bf The lower bound.} From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $\al^*\ge \al$ if $\al$ satisfies
\be\label{solval2}
\begin{split}
&(1-\oo(1)) \cdot \sum_{i=1}^2 pd^2 \frac{(\al\rho_i)^2\cdot \al (\rho_1+\rho_2)}{[\al ( \rho_1+\rho_2) - 1]^3}  \bigbrace{1 - \sqrt{\frac{1}{\al \rho_i}}}^4 \\
&\ge \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{\al (\rho_1+\rho_2) - 1} . 
\end{split}
\ee
We then follow similar arguments as the above proof for the upper bound.

In order to solve \eqref{solval2}, we consider the case $\min(\rho_1,\rho_2)\ge 200$. With some basic calculations, one can show that the sum on the left-hand side of \eqref{solval2} satisfies 
where $\rho:=\rho_1+\rho_2$. 
$$ \frac{1}{[1 - (\al \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho_1+\rho_2}} - \sqrt{\frac{1}{\al \rho}}}^4 >  0.33.$$
Thus the following inequality implies \eqref{solval2}:
\be\label{solval2add}
\begin{split}
&0.33 pd^2 > \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{\al (\rho_1+\rho_2) - 1} .
\end{split}
\ee
There are two cases: if 
\be\nonumber
\begin{split}
&0.33 pd^2 \ge \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1},
\end{split}
\ee
then we always have negative transfer for all choice of $0\le \al \le 1$; otherwise, we can solve from \eqref{solval2add} the following lower bound on $\al^*$:
\begin{align*}
\al^* &>  \frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  - (\rho_1-1)(\rho_2-1) \cdot 0.33 \frac{pd^2}{\sigma^2}}+ 1\right] \\
& >\frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1}+\frac1{\rho_2}  - \frac14\frac{pd^2}{\sigma^2}}+ 1\right] .
\end{align*}
This concludes the proof.

We remark that the constants 1.32 and 0.33 are chosen for all values of $\rho_1/\rho_2$. For a fixed ratio $\rho_1/\rho_2$, one can probably choose much better constants.
\end{proof}
